---
title: "p8105_hw3_kl3181"
author: Kelley Lou
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##Problem 1
Load in dataset.
```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles and which are most items from?
```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Making a plot for number of items ordered in each aisle. 
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 1000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Making a plot to show the three most popular items. 
```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "pacakaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Making a table to compare mean hour of the day at which apples and ice cream are ordered.
```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

##Problem 2
Load in the dataset.
```{r}
accel_df = 
  read_csv(
    "./accel_data.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    weekday = if_else(day %in% c("Saturday","Sunday"), "Weekend", "Weekday"),
  ) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    values_to = "activity_count"
  ) %>% 
```
activity count and minute variables, add new variable (saturday and sunday are weekend- mutate), 
encode with variable classes (activity count - number, day - character or factor, is order correct?)
is minute of the day numerical? SHOULD NOT BE CHARACTER

total activity variable = 
group_by(week, day)
summarize (mean activity count)
35 days - total activity count
make the table easier to read (week number, day of week, activity count)
dow as a column - use a factor to make sure the order is not alphabetical

single-panel plot = 
ggplot
24 hr of activity, what does it look like at each minute of each day
  min = x, activity count = y
geom_line
aes color= dow

##Problem 3
```{r}
data("ny_noaa")
```
separate to get three variables year month and day
check in NY NOAA dataset description to see if the unit is reasonable for temp, precipitation, snowfall - convert to degrees (not tenths)
snowfall values (count) 

two panel plot = 
data manipulation
  filter jan and july
  group by station, year, month
  summarize average max temp
ggplot 
  plot time course for station 1, 2, 3, 4, 5
  average max temp over years, separately for each station (TWO PANEL)
  faceting for two panel
  global warming? are some stations always colder than others?
  
two panel plot = 
  two distinct plots, then join them (patchwork)
  tmax vs. tmin
    hex plot 
  snowfall > 0, <100
    filter 
    box plot, violin plot, ridge plot
    one box for each year
  